---
title: "Site Unit Summary"
format: 
  html:
    theme: journal
    toc: true
params:
  plot_number: "00000"
  db_path: "../data/vpro.duckdb"
  parquet_dir: ""
---

```{r setup, include=FALSE}
library(duckdb)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)

options(knitr.table.format = "html")

# Connect to DB (use parquet snapshot when provided)
parquet_dir <- params$parquet_dir
if (!is.null(parquet_dir) && nzchar(trimws(parquet_dir))) {
  con <- dbConnect(duckdb(), dbdir = ":memory:")
  tbls <- c("Sample_Env", "vw_USysAllVeg", "Sample_Humus", "Sample_Mineral")
  for (tbl in tbls) {
    file_name <- paste0(gsub("[^A-Za-z0-9_]", "_", tbl), ".parquet")
    file_path <- file.path(parquet_dir, file_name)
    if (file.exists(file_path)) {
      dbExecute(
        con,
        sprintf("CREATE TABLE %s AS SELECT * FROM read_parquet('%s')", tbl, gsub("'", "''", file_path))
      )
    }
  }
} else {
  con <- dbConnect(duckdb(), params$db_path, read_only = TRUE)
}
```

```{r load-data, include=FALSE}
# 1. Header Data
query_env <- "SELECT * FROM Sample_Env WHERE plotnumber = ?"
env_data <- dbGetQuery(con, query_env, list(params$plot_number))

normalize_env_cols <- function(df) {
  if (nrow(df) == 0) return(df)
  cols <- names(df)
  pick_col <- function(candidates) {
    idx <- which(tolower(cols) %in% tolower(candidates))
    if (length(idx) > 0) cols[[idx[1]]] else NA_character_
  }
  map <- list(
    plotnumber = c("plotnumber", "PlotNumber"),
    projectid = c("projectid", "ProjectID"),
    date = c("date", "Date"),
    sitesurveyor = c("sitesurveyor", "SiteSurveyor"),
    `_location` = c("_location", "location", "Location"),
    latitude = c("latitude", "Latitude"),
    longitude = c("longitude", "Longitude"),
    elevation = c("elevation", "Elevation"),
    aspect = c("aspect", "Aspect"),
    utmzone = c("utmzone", "UTMZone"),
    utmeasting = c("utmeasting", "UTMEasting"),
    utmnorthing = c("utmnorthing", "UTMNorthing"),
    slopegradient = c("slopegradient", "SlopeGradient"),
    ntsmapsheet = c("ntsmapsheet", "NtsMapSheet"),
    mesoslopeposition = c("mesoslopeposition", "MesoSlopePosition"),
    surfaceshape = c("surfaceshape", "SurfaceShape"),
    moistureregime = c("moistureregime", "MoistureRegime"),
    nutrientregime = c("nutrientregime", "NutrientRegime"),
    sitenotes = c("sitenotes", "SiteNotes")
  )
  for (target in names(map)) {
    source <- pick_col(map[[target]])
    if (!is.na(source)) df[[target]] <- df[[source]]
  }
  df
}

env_data <- normalize_env_cols(env_data)

# 2. Vegetation Data
query_veg <- "SELECT * FROM vw_USysAllVeg WHERE PlotNumber = ? ORDER BY MyLayer, Species"
veg_data <- dbGetQuery(con, query_veg, list(params$plot_number))

normalize_veg_cols <- function(df) {
  if (nrow(df) == 0) return(df)
  cols <- names(df)
  pick_col <- function(candidates) {
    idx <- which(tolower(cols) %in% tolower(candidates))
    if (length(idx) > 0) cols[[idx[1]]] else NA_character_
  }
  species_col <- pick_col(c("species", "species_code", "Species"))
  cover_col <- pick_col(c("cover", "cover_value", "Cover"))
  layer_col <- pick_col(c("mylayer", "layer", "MyLayer"))
  if (!is.na(species_col)) df$Species <- df[[species_col]]
  if (!is.na(cover_col)) df$Cover <- df[[cover_col]]
  if (!is.na(layer_col)) df$MyLayer <- df[[layer_col]]
  df
}

veg_data <- normalize_veg_cols(veg_data)

# 3. Soil Data
humus_data <- dbGetQuery(con, "SELECT * FROM Sample_Humus WHERE plotnumber = ? ORDER BY horizon", list(params$plot_number))
mineral_data <- dbGetQuery(con, "SELECT * FROM Sample_Mineral WHERE plotnumber = ? ORDER BY horizon", list(params$plot_number))

dbDisconnect(con, shutdown = TRUE)
```

# Plot: `r params$plot_number`

**Project:** `r if(nrow(env_data)>0) env_data$projectid[1] else "N/A"`  
**Date:** `r if(nrow(env_data)>0) env_data$date[1] else "N/A"`  
**Surveyor:** `r if(nrow(env_data)>0) env_data$sitesurveyor[1] else "N/A"`  

## Location
```{r location, echo=FALSE}
if(nrow(env_data) > 0) {
  loc_df <- data.frame(
    Attribute1 = c("Location", "Lat", "Long", "Elev (m)", "Aspect"),
    Value1 = c(env_data[["_location"]][1], env_data$latitude[1], env_data$longitude[1], env_data$elevation[1], env_data$aspect[1]),
    Attribute2 = c("UTM Zone", "Easting", "Northing", "Slope (%)", "Map Sheet"),
    Value2 = c(env_data$utmzone[1], env_data$utmeasting[1], env_data$utmnorthing[1], env_data$slopegradient[1], env_data$ntsmapsheet[1])
  )
  
  kbl(loc_df, col.names = NULL) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
} else {
  cat("No environment data found.")
}
```

## Site Description

**Meso Slope:** `r if(nrow(env_data)>0) env_data$mesoslopeposition[1] else "N/A"`  
**Surface Shape:** `r if(nrow(env_data)>0) env_data$surfaceshape[1] else "N/A"`  
**Moisture Regime:** `r if(nrow(env_data)>0) env_data$moistureregime[1] else "N/A"`  
**Nutrient Regime:** `r if(nrow(env_data)>0) env_data$nutrientregime[1] else "N/A"`  

**Notes:**  
<div class="alert alert-info">
`r if(nrow(env_data)>0 && !is.na(env_data$sitenotes[1])) env_data$sitenotes[1] else "No notes recorded."`
</div>

## Vegetation

```{r veg-table, echo=FALSE, results='asis'}
if (nrow(veg_data) > 0) {
  
  # Map Layer Codes to Names
  layer_names <- c(
    "1" = "Layer A1 (Dominant Trees)",
    "2" = "Layer A2 (Codominant Trees)",
    "3" = "Layer A3 (Intermediate Trees)", 
    "4" = "Layer B1 (Tall Shrubs)",
    "5" = "Layer B2 (Low Shrubs)",
    "6" = "Layer C (Herbs)",
    "7" = "Layer D (Moss/Lichen)"
  )
  
  veg_data$LayerName <- layer_names[veg_data$MyLayer]
  veg_data$LayerName[is.na(veg_data$LayerName)] <- paste("Layer", veg_data$MyLayer[is.na(veg_data$LayerName)])
  
  veg_display <- veg_data %>%
    select(LayerName, Species, Cover)
    
  veg_split <- split(veg_display, veg_display$LayerName)
  
  for (layer in names(veg_split)) {
    cat(paste0("\n### ", layer, "\n"))
    
    print(kbl(veg_split[[layer]] %>% select(-LayerName), col.names = c("Species", "Cover (%)")) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "left"))
        
    cat("\n")
  }
    
} else {
  cat("No vegetation data recorded.")
}
```

## Soil Profile

### Humus Horizons
```{r humus, echo=FALSE}
if (nrow(humus_data) > 0) {
  humus_data %>%
    select(horizon, upperdepth, lowerdepth, humusstructuredegree, humusstructurekind, humusformph) %>%
    kbl(col.names = c("Horizon", "Upper", "Lower", "Struct Deg", "Struct Kind", "pH")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = T)
} else {
  cat("No humus data.")
}
```

### Mineral Horizons
```{r mineral, echo=FALSE}
if (nrow(mineral_data) > 0) {
  mineral_data %>%
    select(horizon, upperdepth, lowerdepth, texture, percentcoarsefragstotal, colour) %>%
    kbl(col.names = c("Horizon", "Upper", "Lower", "Texture", "% Coarse", "Colour")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = T)
} else {
  cat("No mineral data.")
}
```
