---
title: "Site Unit Summary"
output: pdf_document
params:
  plot_number: "00000"
  db_path: "../data/vpro.duckdb"
---

```{r setup, include=FALSE}
library(duckdb)
library(dplyr)
library(tidyr)
library(knitr)

# Connect to DB (Read Only to avoid locks from main app)
# If knitting from RStudio, getwd() is the Rmd dir.
# If running via shiny, it depends.
# We passed db_path param to help.
con <- dbConnect(duckdb(), params$db_path, read_only = TRUE)
```

```{r load-data, include=FALSE}

# 1. Header Data
# using vw_USysEnv or just Sample_Env
query_env <- "SELECT * FROM Sample_Env WHERE plotnumber = ?"
env_data <- dbGetQuery(con, query_env, list(params$plot_number))

# 2. Vegetation Data
# We use the View vw_USysAllVeg
query_veg <- "SELECT * FROM vw_USysAllVeg WHERE PlotNumber = ? ORDER BY MyLayer, Species"
veg_data <- dbGetQuery(con, query_veg, list(params$plot_number))

# 3. Soil Data
humus_data <- dbGetQuery(con, "SELECT * FROM Sample_Humus WHERE plotnumber = ? ORDER BY horizon", list(params$plot_number))
mineral_data <- dbGetQuery(con, "SELECT * FROM Sample_Mineral WHERE plotnumber = ? ORDER BY horizon", list(params$plot_number))

dbDisconnect(con)
```

# Plot: `r params$plot_number`

**Project:** `r env_data$projectid[1]`  
**Date:** `r env_data$date[1]`  
**Surveyor:** `r env_data$sitesurveyor[1]`  

## Location
| Attribute | Value | Attribute | Value |
|-----------|-------|-----------|-------|
| **Location** | `r env_data[["_location"]][1]` | **UTM Zone** | `r env_data$utmzone[1]` |
| **Lat** | `r env_data$latitude[1]` | **Easting** | `r env_data$utmeasting[1]` |
| **Long** | `r env_data$longitude[1]` | **Northing** | `r env_data$utmnorthing[1]` |
| **Elev (m)** | `r env_data$elevation[1]` | **Slope (%)** | `r env_data$slopegradient[1]` |
| **Aspect** | `r env_data$aspect[1]` | **Map Sheet** | `r env_data$ntsmapsheet[1]` |

## Site Description
**Meso Slope:** `r env_data$mesoslopeposition[1]`  
**Surface Shape:** `r env_data$surfaceshape[1]`  
**Moisture Regime:** `r env_data$moistureregime[1]`  
**Nutrient Regime:** `r env_data$nutrientregime[1]`  

**Notes:**  
`r env_data$sitenotes[1]`

## Vegetation

```{r veg-table, echo=FALSE, warning=FALSE}

if (nrow(veg_data) > 0) {
  # Pivot nice table: Layers as Columns? Or just list?
  # Standard VPro report lists them by Layer.
  
  # Map Layer Codes to Names
  layer_names <- c(
    "1" = "Layer A1 (Dominant Trees)",
    "2" = "Layer A2 (Codominant Trees)",
    "3" = "Layer A3 (Intermediate Trees)", 
    "4" = "Layer B1 (Tall Shrubs)",
    "5" = "Layer B2 (Low Shrubs)",
    "6" = "Layer C (Herbs)",
    "7" = "Layer D (Moss/Lichen)"
  )
  
  veg_data$LayerName <- layer_names[veg_data$MyLayer]
  veg_data$LayerName[is.na(veg_data$LayerName)] <- paste("Layer", veg_data$MyLayer)
  
  veg_display <- veg_data %>%
    select(LayerName, Species, Cover)
    
  # knitr::kable(veg_display)
  
  veg_split <- split(veg_display, veg_display$LayerName)
  
  for (layer in names(veg_split)) {
    cat(paste0("\n**", layer, "**\n"))
    print(kable(veg_split[[layer]] %>% select(-LayerName), col.names = c("Species", "Cover")))
    cat("\n")
  }
    
} else {
  cat("No vegetation data recorded.")
}
```

## Soil Profile

### Humus Horizons
```{r humus, echo=FALSE}
if (nrow(humus_data) > 0) {
  humus_data %>%
    select(horizon, upperdepth, lowerdepth, humusstructuredegree, humusstructurekind, humusformph) %>%
    kable(col.names = c("Horizon", "Upper", "Lower", "Struct Deg", "Struct Kind", "pH"))
} else {
  cat("No humus data.")
}
```

### Mineral Horizons
```{r mineral, echo=FALSE}
if (nrow(mineral_data) > 0) {
  mineral_data %>%
    select(horizon, upperdepth, lowerdepth, texture, percentcoarsefragstotal, colour) %>%
    kable(col.names = c("Horizon", "Upper", "Lower", "Texture", "% Coarse", "Colour"))
} else {
  cat("No mineral data.")
}
```
