---
title: "Hierarchy Collapsible Tree"
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
params:
  db_path: "../data/vpro.duckdb"
  collapsed: true
  report_title: "Hierarchy collapsible tree"
---

```{r setup}
library(DBI)
library(duckdb)
library(dplyr)
library(tibble)
library(knitr)
library(collapsibleTree)
library(data.tree)
library(htmltools)

resolve_db_path <- function(db_path) {
  input_dir <- tryCatch(dirname(knitr::current_input(dir = TRUE)), error = function(e) getwd())
  candidates <- unique(c(
    db_path,
    file.path(input_dir, db_path),
    file.path(getwd(), db_path),
    file.path(input_dir, "../data/vpro.duckdb"),
    file.path(getwd(), "data/vpro.duckdb")
  ))
  candidates <- candidates[file.exists(candidates)]
  if (length(candidates) == 0) {
    stop("Unable to locate the VPRO DuckDB file.")
  }
  normalizePath(candidates[[1]], winslash = "/", mustWork = TRUE)
}

connect_hierarchy_db <- function(db_path) {
  db_path <- resolve_db_path(db_path)
  direct <- tryCatch(
    DBI::dbConnect(duckdb::duckdb(), db_path, read_only = TRUE),
    error = function(e) e
  )

  if (!inherits(direct, "error")) {
    return(list(con = direct, source = "direct", db_path = db_path))
  }

  temp_copy <- file.path(tempdir(), paste0("vpro-hierarchy-", Sys.getpid(), "-", as.integer(Sys.time()), ".duckdb"))
  copied <- file.copy(db_path, temp_copy, overwrite = TRUE)
  if (!isTRUE(copied)) {
    stop("Could not copy the locked DuckDB file for read-only reporting.")
  }

  con <- DBI::dbConnect(duckdb::duckdb(), temp_copy, read_only = TRUE)
  list(con = con, source = "temp-copy", db_path = db_path)
}

normalize_hierarchy <- function(df) {
  cols <- names(df)
  pick_col <- function(candidates) {
    idx <- which(tolower(cols) %in% tolower(candidates))
    if (length(idx) == 0) return(NA_character_)
    cols[[idx[[1]]]]
  }

  id_col <- pick_col(c("id", "ID"))
  name_col <- pick_col(c("name", "Name", "_name"))
  parent_col <- pick_col(c("parent", "Parent"))
  level_col <- pick_col(c("level", "Level", "_level"))
  order_col <- pick_col(c("myorder", "MyOrder"))

  tibble(
    ID = as.numeric(df[[id_col]]),
    Name = as.character(df[[name_col]]),
    Parent = as.numeric(df[[parent_col]]),
    Level = as.integer(df[[level_col]]),
    MyOrder = if (!is.na(order_col)) as.character(df[[order_col]]) else NA_character_
  ) |>
    filter(!is.na(ID), nzchar(Name)) |>
    arrange(Level, MyOrder, Name) |>
    distinct(ID, .keep_all = TRUE)
}

shorten_label <- function(label, max_chars = 26) {
  ifelse(
    is.na(label) | nchar(label) <= max_chars,
    label,
    paste0(substr(label, 1, max_chars - 3), "...")
  )
}

make_display_labels <- function(labels, ids, max_chars = 26) {
  short <- shorten_label(labels, max_chars = max_chars)
  dup_count <- ave(short, short, FUN = length)
  ifelse(dup_count > 1, paste0(short, " #", ids), short)
}

fetch_hierarchy <- function(db_path) {
  db_info <- connect_hierarchy_db(db_path)
  on.exit({
    try(DBI::dbDisconnect(db_info$con, shutdown = TRUE), silent = TRUE)
  }, add = TRUE)

  if (!DBI::dbExistsTable(db_info$con, "Hierarchy")) {
    stop("Hierarchy table was not found in the DuckDB database.")
  }

  hier <- DBI::dbGetQuery(db_info$con, "SELECT * FROM Hierarchy")
  attr(hier, "db_source") <- db_info$source
  attr(hier, "db_path") <- db_info$db_path
  hier
}

build_hierarchy_paths <- function(nodes) {
  child_lookup <- split(nodes$ID, ifelse(is.na(nodes$Parent), "ROOT", as.character(nodes$Parent)))
  child_lookup <- lapply(child_lookup, function(values) {
    ordered <- nodes |>
      filter(ID %in% values) |>
      arrange(Level, MyOrder, Name)
    ordered$ID
  })
  name_lookup <- setNames(nodes$Name, nodes$ID)

  roots <- nodes$ID[is.na(nodes$Parent) | !(nodes$Parent %in% nodes$ID)]
  roots <- roots[order(match(roots, nodes$ID))]

  walk_paths <- function(node_id, lineage = character(), visited = numeric()) {
    if (node_id %in% visited) {
      return(list(list(ID = node_id, path = c(lineage, paste0(name_lookup[[as.character(node_id)]], " [cycle]")))))
    }

    current_path <- c(lineage, name_lookup[[as.character(node_id)]])
    children <- child_lookup[[as.character(node_id)]]
    if (is.null(children) || length(children) == 0) {
      return(list(list(ID = node_id, path = current_path)))
    }

    out <- list()
    for (child_id in children) {
      out <- c(out, walk_paths(child_id, current_path, c(visited, node_id)))
    }
    out
  }

  paths <- list()
  for (root_id in roots) {
    paths <- c(paths, walk_paths(root_id))
  }
  paths
}

build_collapsible_tree_node <- function(nodes, root_label = "Hierarchy") {
  child_lookup <- split(nodes$ID, ifelse(is.na(nodes$Parent), "ROOT", as.character(nodes$Parent)))
  child_lookup <- lapply(child_lookup, function(values) {
    ordered <- nodes |>
      filter(ID %in% values) |>
      arrange(Level, MyOrder, Name)
    ordered$ID
  })
  row_lookup <- split(nodes, nodes$ID)

  root <- data.tree::Node$new(root_label)
  root$tooltip_html <- "<strong>Hierarchy</strong><br>Root node"
  root$isCollapsed <- FALSE

  walk_node <- function(node_id, parent_node, lineage = character(), visited = numeric()) {
    row <- row_lookup[[as.character(node_id)]][1, ]
    current_path <- c(lineage, row$FullName)

    child <- parent_node$AddChild(row$DisplayName)
    child$tooltip_html <- paste0(
      "<strong>", htmltools::htmlEscape(row$FullName), "</strong>",
      "<br>ID: ", row$ID,
      "<br>Level: ", row$Level,
      "<br>Children: ", row$child_count,
      "<br>Path: ", htmltools::htmlEscape(paste(current_path, collapse = " / "))
    )
    child$isCollapsed <- row$child_count > 0

    if (node_id %in% visited) {
      child$tooltip_html <- paste0(child$tooltip_html, "<br>Cycle detected")
      return(invisible(NULL))
    }

    children <- child_lookup[[as.character(node_id)]]
    if (is.null(children) || length(children) == 0) {
      return(invisible(NULL))
    }

    for (child_id in children) {
      walk_node(child_id, child, current_path, c(visited, node_id))
    }
  }

  roots <- nodes$ID[is.na(nodes$Parent) | !(nodes$Parent %in% nodes$ID)]
  roots <- roots[order(match(roots, nodes$ID))]

  for (root_id in roots) {
    walk_node(root_id, root)
  }

  root
}

hier_raw <- fetch_hierarchy(params$db_path)
hier <- normalize_hierarchy(hier_raw)
child_counts <- hier |>
  filter(!is.na(Parent)) |>
  count(Parent, name = "child_count")

nodes <- hier |>
  left_join(child_counts, by = c("ID" = "Parent")) |>
  mutate(
    child_count = ifelse(is.na(child_count), 0L, child_count),
    FullName = Name,
    DisplayName = make_display_labels(Name, ID, max_chars = 24)
  )

collapsible_tree_node <- build_collapsible_tree_node(nodes)

summary_tbl <- tibble(
  metric = c("Node count", "Root count", "Max level", "Database source"),
  value = c(
    nrow(nodes),
    sum(is.na(nodes$Parent) | !(nodes$Parent %in% nodes$ID)),
    max(nodes$Level, na.rm = TRUE),
    attr(hier_raw, "db_source")
  )
)
```

# `r params$report_title`

This is a focused example using only `collapsibleTree`. It starts collapsed by default so the user can uncover the hierarchy one branch at a time, with a direct click on the node they want to inspect.

## Data Summary

```{r summary}
knitr::kable(summary_tbl, align = c("l", "l"))
```

## Collapsible Tree Example

```{r tree}
collapsibleTree::collapsibleTree(
  collapsible_tree_node,
  tooltip = TRUE,
  tooltipHtml = "tooltip_html",
  collapsed = "isCollapsed",
  linkLength = 220,
  fontSize = 14,
  zoomable = TRUE,
  width = "100%",
  height = 1200
)
```

## Notes

- Nodes can be expanded and collapsed by clicking directly on the node label.
- Branches now use explicit collapse flags, so the tree opens folded by default and reveals one path at a time.
- Long labels are shortened on the canvas and the full label is shown in the tooltip.
- This widget is better for exploration than for publication-style graphics.